Lactate transport inhibition therapeutically reprograms fibroblast

### 2024-02-15 manuscript/manuscript.qmd
finish up ROS data

Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which couple lactate oxidation to NADH production. Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression, we next examined the impact of lactate transport inhibition on cellular redox homeostasis. Consistent with an increase in intracellular lactate, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A-C**).

>>>>

Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which transfer electrons from lactate to NADH. Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression [@jain2013, @schworer2020], we next examined the impact of lactate transport inhibition on cellular redox homeostasis. Consistent with an increase in intracellular lactate with MCT4 inhibition, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A**), which was associated with a reduction in total ROS as measured by the CellROX fluorescent probe (**Figure 7B**). By contrast, the NADPH/NADP^+^ was unchanged by either TGFβ or MCT inhibitors (**Supplemental Figure 7A**). Notably, we did not observe TGFβ-dependent increases in total ROS production nor an impact of TGFβ or inhibitors on mitochondrial superoxide production (**Supplemental Figure 7B**).

Proline biosynthesis has been previously suggested as a mechanism by which NIH-3T3 fibroblasts prevent excessive ROS accumulation following TGFβ stimulation [@schworer2020]. In this model, proline synthesis from glutamine consumes reducing equivalents from NADPH and NADH, thereby ameliorating reductive stress and decreasing ROS production. Compared to this prior study, we did not find that TGFβ increased proline in our primary lung fibroblasts (**Supplemental Figure 7C**). Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest (~10%) compared to the 40% previously reported (**Supplemental Figure 7D**). While we observed trends toward increased proline production from glutamine with MCT4 inhibition, this mechanism seems unlikely to account for the effects of MCT inhibitors on ROS generation.
